The Space of Feasible Execution Times for Asynchronous Periodic Task Systems using Definitive Idle Times

### 2013-09-24 dit-paper_ieee.tex
size ratio n systems
Reworked conclusion

\begin{abstract}
	Sensitivity analysis for real-time systems provides efficient methods to determine feasibility
	conditions in the case of small changes in task parameters, such as
	different WCET values when porting a system to a different platform. The \mbox{C-space} describes the set of execution
	times for which a system is feasible, which allows system designers to quickly
	find the best platform for a given task set. This paper extends previous works
	generalizing the description of the C-space for Earliest Deadline First
	scheduling to systems with initial task offsets. The feasibility gain offered
	by offsets is then expressed as the volume ratio between the C-space of an
	asynchronous system and its corresponding synchronous system. The problem
	requires solving multiple similar instances of the Chinese Remainder Problem,
	for which an algorithm efficient in this context is provided.
\end{abstract}

@@

	Finally, in section~\ref{sct:expCspaceGain}, we use a simulation to quantify the feasibility
	gain of asynchronous systems in terms of volume of the C-space.

@@

	\subsection{Model}
		We consider a discrete timeline and tasks systems comprised of periodic
		tasks with constrained deadlines on uniprocessor platforms. Each task
		$\tau_i$ is represented by a tuple $(O_i, C_i, D_i, T_i)$ where $O_i$ is the offset value,
		$C_i$ the WCET, $D_i$ the relative deadline and $T_i$ the period, which is the exact time between two consecutive arrivals.
		The constrained deadline property implies that $D_i \leq T_i$. In
		the following, we differentiate between \emph{synchronous systems} (where $O_i
		= 0$ for all tasks) and \emph{asynchronous systems} (which are not synchronous).

@@

		Furthermore, the following notations are used:
		\begin{itemize}
			\item $\tau = \{\tau_1, \cdots, \tau_{n}\}$ is called the task system
			\item $H = \operatorname{lcm}(T_1, \cdots, T_{n})$
			\item $O_{max} = \max (O_1, \cdots, O_{n})$
			\item An \emph{idle time} is an instant $t$ at which each job that arrived
			strictly before $t$ has finished its execution.
			\item A \emph{busy period} is a time interval between two consecutive idle times in
			which at least one job is executing.
			\item The interval between the initial time ($t=0$) and the first
			idle time after $t=0$ is called the \emph{first busy period}.
		\end{itemize}

@@

		The following necessary and sufficient condition of feasibility
		based on the DBF is given in \cite{baruah1999generalized}:
		\begin{theorem}
			\begin{equation*}
				\begin{array}{c}
					\tau \: \text{is feasible}  \iff
					\dbf{t_1, t_2} \leqslant t_2 - t_1 \; \forall t_1, t_2 \mid  0 \leq t_1 \leq t_2
				\end{array}
			\end{equation*}
		\end{theorem}
        It is easy to see that the condition is at least necessary, as if there exists values of $t_1$ and $t_2$ such that $\dbf{t_1, t_2} > t_2 - t_1$, it means that more than $t_2 - t_1$ units of computation have to be done in $t_2 - t_1$ units of time. Therefore the system would not be feasible.

@@

      However, this description contains an infinite number of constraints, most of
      which are redundant. While this does not harm the precision of the
      test, a finite description is needed in practice. The description should also be as small as possible as this can greatly
      reduce the computation time of practical feasibility tests. In Section~\ref{sct:removeRedundancy} a method to reduce the size of this set of constraints is given.

@@

  Even then, the number of constraints to consider is exponential in the number
  of tasks. While it has been shown that constraints that cover intervals ending
  after the end of the first busy period are also redundant, this value depends on the $C_i$ and is
  therefore not usable for the purposes of this paper. The
  following concept, initially described by \cite{lipariaverage}, is thus introduced. It gives an upper bound for the end of the first busy period and does not depend on the execution times.

@@

		\begin{remark} (from the definitions):
			\begin{itemize}
				\item Contrary to other idle times, DITs are independent of the
					scheduling or the execution times of the jobs.
				\item In a feasible system, every DIT is an idle time.
				\item In a feasible system, the first DIT happens after the end of the
					first busy period.
				\item In synchronous systems any $kH$ with $k \geqslant 0$ is a DIT and we
					also have $t_d \leqslant H$.
			\end{itemize}
		\end{remark}

		The value of the first DIT can be found by solving the following modular
		equations system:
		\begin{equation}
			\begin{split}
				 t_d = \min(& \{ t \; s.t. \; t > 0 \; and \\
						& \forall \tau_i \in \tau, \; \exists a_i \in [D_i, T_i]: t \equiv a_i \;
						(mod \; T_i) \})
			\end{split}
		\end{equation}
  The authors of \cite{george2009characterization} give a method to solve this system for cases in which periods are pairwise coprime.

@@

	We thus checked the existence condition on 10000 systems generated by:
	\begin{itemize}
		\item Generating utilization values with the \texttt{UUniFast} algorithm described in \cite{bini2005measuring}, with:
		\begin{itemize}
			\item System utilization chosen uniformly between 0.25 and 0.75
			\item The number of tasks set to several values (see Fig.~\ref{fig:noFPDIT})
		\end{itemize}
		\item Choosing the $T_i$ uniformly between 5 and 20 (in order to have a manageable value of $H$), which also gives us the $C_i = \lfloor U_i \, T_i \rfloor$
		\item Choosing the $O_i$ with a normal distribution of mean $T_{min}$ and standard deviation $\frac{T_{max} - T_{min}}{2}$
		\item Choosing the $D_i$ uniformly in the interval $[T_i - CDF(T_i - C_i), T_i]$, where $CDF$ varies between 0 and 1 and is called the \emph{Constrained Deadline Factor}. It allows us to generate systems where the deadlines are closer to the period.
	\end{itemize}

@@

@@

In this paper, we presented a general approach to define the C-space of a constrained periodic task set for EDF on uniform platform. Possible future works includes an extension to the study of systems with preemption costs or running on multiprocessor platforms, as the value of the DIT does not depend on preemptions or the number of processors.

Future works could also repeat the simulation of section~\ref{sct:expCspaceGain} with the algorithm described in \cite{grenier2008pushing}, which determine offset values with optimal feasibility. The
feasibility gain could then be expressed using the C-space.

>>>>

\begin{abstract}
	Sensitivity analysis for real-time systems provides efficient methods to determine feasibility
	conditions in the case of small changes in task parameters, such as
	different WCET values when porting a system to a different platform. The \mbox{C-space} describes the set of execution
	times for which a system is feasible, which allows system designers to quickly
	find the best platform for a given task set. This paper extends previous works
	to generalize the description of the C-space for Earliest Deadline First
	scheduling to systems with initial task offsets. The feasibility gain offered
	by offsets is then expressed as the volume ratio between the C-space of an
	asynchronous system and its corresponding synchronous system. The problem
	requires solving multiple similar instances of the Chinese Remainder Problem,
	for which an algorithm efficient in this context is provided.
\end{abstract}

@@

	Finally, in section~\ref{sct:expCspaceGain}, we use a simulation to quantify the feasibility
	gain of asynchronous system in terms of volume of the C-space.

@@

	\subsection{Model}
		We consider a discrete timeline and tasks systems $\tau$ comprised of periodic
		tasks with constrained deadlines on uniprocessor platforms. Each task
		$\tau_i$ is represented by a tuple $(O_i, C_i, D_i, T_i)$ where $O_i$ is the offset value,
		$C_i$ the WCET, $D_i$ the relative deadline and $T_i$ the period, which is the exact time between two consecutive arrivals.
		The constrained deadline property implies that $D_i \leq T_i$. In
		the following, we differentiate between \emph{synchronous systems} (where $O_i
		= 0$ for all tasks) and \emph{asynchronous systems} (which are not synchronous).

@@

		Furthermore, the following notations are used:
		\begin{itemize}
			\item $\tau = \{\tau_1, \cdots, \tau_{n}\}$
			\item $H = \operatorname{lcm}(T_1, \cdots, T_{n})$
			\item $O_{max} = \max (O_1, \cdots, O_{n})$
			\item An \emph{idle time} is an instant $t$ at which each job that arrived
			strictly before $t$ has finished its execution.
			\item A \emph{busy period} is a time interval between two consecutive idle times in
			which at least one job is executing.
			\item The interval between the initial time ($t=0$) and the first
			idle time after $t=0$ is called the \emph{first busy period}.
		\end{itemize}

@@

		The following necessary and sufficient condition of feasibility
		based on the DBF is given in \cite{baruah1999generalized}:
		\begin{theorem}
			\begin{equation*}
				\begin{array}{c}
					\tau \: \text{is feasible}  \iff
					\dbf{t_1, t_2} \leqslant t_2 - t_1 \; \forall t_1, t_2 \mid  0 \leq t_1 \leq t_2
				\end{array}
			\end{equation*}
		\end{theorem}
        It is easy to see that the condition is at least necessary, as if there exists values of $t_1$ and $t_2$ such that $\dbf{t_1, t_2} > t_2 - t_1$, it means that more than $t_2 - t_1$ unit of computation have to be done in $t_2 - t_1$ unit of time. Therefore the system would be unfeasible.

@@

      However, this description contains an infinite number of constraints, most of
      which are redundant. While this does not harm the precision of the
      test, a smaller description is preferred as it can greatly
      reduce the computation time of practical feasibility tests. In Section~\ref{sct:removeRedundancy} a method to reduce the size of this set of constraints is given.

@@

  Even then, the number of constraints to consider is exponential in the number
  of tasks. While it has been shown that constraints that cover intervals ending
  after the end of the first busy period are also redundant, this value depends on the $C_i$ and is
  therefore not usable for the purposes of this paper. The
  following concept, initially described by \cite{lipariaverage}, is thus introduced. It gives an upper bound for to the end of the busy period and does not depend on the execution times.

@@

		\begin{remark} (from the definitions):
			\begin{itemize}
				\item Contrary to other idle times, DITs are independent of the
					scheduling or the execution times of the jobs.
				\item In a feasible system, every DIT is an idle time.
				\item In a feasible system, the first DIT happens after the end of the
					first busy period.
				\item In synchronous systems any $kH$ with $k \geqslant 0$ is a DIT and we
					also have at least $t_d \leqslant H$.
			\end{itemize}
		\end{remark}

		The value of the first DIT can be found by solving the following modular
		equation system:
		\begin{equation}
			\begin{split}
				 t_d = \min(& \{ t \; s.t. \; t > 0 \; and \\
						& \forall \tau_i \in \tau, \; \exists a_i \in [D_i, T_i]: t \equiv a_i \;
						(mod \; T_i) \})
			\end{split}
		\end{equation}
  The authors of \cite{george2009characterization} give a method to solve this system for cases in which periods are pairwise coprime.

@@

	We thus checked the existence condition on 10000 systems generated by:
	\begin{itemize}
		\item Generating utilization values with the \texttt{UUniFast} algorithm described in \cite{bini2005measuring}, with:
		\begin{itemize}
			\item The system utilization chosen uniformly between 0.25 and 0.75
			\item The number of tasks set to several values (see Fig.~\ref{fig:noFPDIT})
		\end{itemize}
		\item Choosing the $T_i$ uniformly between 5 and 20 (in order to have a manageable value of $H$), which also gives us the $C_i = \lfloor U_i \, T_i \rfloor$
		\item Choosing the $O_i$ with a normal distribution of mean $T_{min}$ and standard deviation $\frac{T_{max} - T_{min}}{2}$
		\item Choosing the $D_i$ uniformly in the interval $[T_i - CDF(T_i - C_i), T_i]$, where $CDF$ varies between 0 and 1 and is called the \emph{Constrained Deadline Factor}. It allows us to generate systems where the deadlines are closer to the period.
	\end{itemize}

@@

    % PRECISER : on ne genere pas de systyemes avec instants synchroes

@@

In this paper, we presented a general approach to concisely define the C-space of a constrained periodic task set for EDF on uniform platforms. We then used the C-space to quantify the feasibility gain given by random offset values, with a final result of about 10~\% more WCET values being feasible. Possible future works includes an extension to the study of systems with preemption costs or running on multiprocessor platforms, as the value of the DIT does not depend on preemptions or the number of processors.

% A mettre quelque part : nombre de contraintes avant/apres dans les expériences (issue #20)

Future works could also study the case of offset-free systems, where the offset values can be chosen. The performance of algorithms such as \cite{grenier2008pushing} and \cite{goossens2003scheduling} (although the latter depends on the execution times), which determine offset values with optimal feasibility, can be expressed using the C-space.
